Fast and Flexible Binary Instrumentation for the Kernel

### 2013-11-27 VEE-14/paper.tex
Fixed up some sub-sub-sections, and described syscall optimizations.

\newsavebox\translatedjcc
\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
cmp $0x0, %rax
edge_jz:
  jz edge_value_is_zero
  ...
edge_value_is_zero:
  call (patch_edge_jz)
\end{lstlisting}
\end{minipage}
\end{lrbox}

\begin{figure}[t!]
\subfloat[Native Code]{\usebox\nativejcc}
\hfill
 \subfloat[Translated Code]{\usebox\translatedjcc}
\caption{\label{fig:direct_edge_code}Example translation of a conditional control flow instruction (\texttt{jz}) and its edge code. The original \texttt{jz} is translated into a \texttt{jz} that uses a 32-bit relative displacement to transfer control to an edge code function call (\texttt{edge\_jz}). When executed, the function call reads its target from a data structure (\texttt{patch\_edge\_jz}), which contains all information necessary to resolve the \texttt{jz}'s target and patch the original \texttt{jz}. Patching is performed with a four byte atomic write to the relative displacement operand of the \texttt{jz} instruction. Finally, the return address of the \texttt{call} instruction is altered to transfer control to the newly resolved target.}
\end{figure}
Like btkernel, Granary translates direct control-flow instructions (CTIs) into equivalent jumps into CTI-specific ``edge code" (\Cref{fig:direct_edge_code}). Edge code lives outside the code cache, so as to not pollute the icache \cite{btkernel}.

@@

\subsubsection{Indirect Control Flow}

@@

Out-edge code is specific to a target basic block, and performs the function of separate chaining in the indirect branch lookup hash table. When control transfers to out-edge code, the out-edge checks to see if the indirect target matches the out-edge code's target basic block. If they match, then the out-edge code restores scratch and flags registers and jumps directly to the target basic block. Otherwise, the out-edge code jumps to another out-edge code or to Granary. If one out-edge code block jumps to another one, then the targets of the two out-edge code blocks collide in the global hash table. In this respect, out-edge code is chained together through control flow, just as colliding entries in a data hash table can be chained together through a linked list.

@@

@@

Granary can target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead. Instrumenting only specific kernel modules is particularly useful for debugging tools because modules represent the bulk of new kernel code under development, and contain the most bugs \cite{FaultsInLinux}.

>>>>

\newsavebox\translatedjcc
\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
  cmp $0x0, %rax
  jz edge_value_is_zero
  jmp edge_fall_through

edge_value_is_zero:
  call *(jz_patch_data)
edge_fall_through:
  call *(jmp_patch_data)
\end{lstlisting}
\end{minipage}
\end{lrbox}

\begin{figure}[t!]
\subfloat[Native Code]{\usebox\nativejcc}
\hfill
\subfloat[Translated Code]{\usebox\translatedjcc}
\caption{\label{fig:direct_edge_code}Example translation of a conditional control flow instruction (\texttt{jz}) and its edge code. The original \texttt{jz} is translated into a \texttt{jz} that uses a 32-bit relative displacement to transfer control to an edge code function call (\texttt{edge\_value\_is\_zero}). When executed, the function call reads its target from an edge data structure (\texttt{jz\_patch\_data}), which contains all information necessary to resolve the \texttt{jz}'s target and patch the original \texttt{jz}. Patching is performed with a four byte atomic write to the relative displacement operand of the \texttt{jz} instruction. Finally, the return address of the \texttt{call} instruction is altered to transfer control to the newly resolved target.}
\end{figure}
Like btkernel, Granary translates direct control-flow instructions (CTIs) into equivalent jumps into CTI-specific ``edge code" (\Cref{fig:direct_edge_code}). Edge code lives outside the code cache, so as to not pollute the icache \cite{btkernel}.

@@

\subsubsection{Indirect Control Flow}\label{sec:ibl}

@@

Out-edge code is specific to a target basic block, and performs the function of separate chaining in the indirect branch lookup hash table. When control transfers to out-edge code, the out-edge checks to see if the indirect target matches the out-edge code's target basic block. If they match, then the out-edge code restores scratch and flags registers and jumps directly to the target basic block. Otherwise, the out-edge code jumps to another out-edge code or to Granary. If one out-edge code block jumps to another one, then the targets of the two out-edge code blocks collide in the global hash table. In this respect, out-edge code is chained together through control flow, just as colliding entries in a typical hash table can be chained together through a linked list.

@@

Granary treats indirect \texttt{jmp}s or \texttt{call}s through the Linux kernel system call table specially. The Linux kernel (as of Linux 3.12) has approximately 300 distinct system call targets, some of which are executed often. To avoid polluting the indirect branch lookup hash table with these targets (\Cref{sec:ibl}), and to avoid the indirect branch lookup overhead, Granary creates a separate, \emph{shadow} system call table.

%re-direct jumps through the system call table

Granary's translation post-processing mechanism identifies indirect control flow through the system call table and alters those CTIs to direct control flow through the shadow system call table instead. The entries of the shadow system call table are ``lazy" entrypoint translator functions. Upon their first execution, these entrypoint translators find or translate the first basic block of the system call handler, and update the shadow system call table accordingly.

%Directing control through the shadow system call table avoids the overhead of going through the indirect branch lookup (IBL) table (\Cref{sec:ibl}), and reduces the number of potential conflicts in the IBL table.

% whose entries are code cache basic block addresses translated according to the starting instrumentation policy.

@@

Unlike DRK and btkernel, Granary can target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead. Instrumenting only specific kernel modules is particularly useful for debugging tools because modules represent the bulk of new kernel code under development, and contain the most bugs \cite{FaultsInLinux}.
